### 2023-07-30 Article_PySF.tex
update(remove_rev_mark) Accepted by CPC. Removed the revision marks.

\begin{abstract}
Tunneling ionization is at the core of strong-field and attosecond science. In this paper, we present \texttt{PyStructureFactor} --- a general \texttt{Python} code towards the calculation of the structure factor in the tunneling ionization rate of common molecules under intense laser fields. The numerical implementation is based on the well-developed weak-field asymptotic theory in the integral representation. The information of the electronic structure of the molecules is obtained via the \texttt{PySCF} quantum chemistry package. \texttt{PyStructureFactor} is a general computational framework that can be utilized to compute the molecular structure factor of various types of molecules, including polar and nonpolar diatomic molecules, degenerate molecules, and \cout{multiplet} \cyan{open-shell} molecules. Examples are given that are benchmarked against known results with good agreements. The present \texttt{PyStructureFactor} is implemented in an efficient manner and is easily applicable towards larger molecules.
\end{abstract}

@@

{\bf \noindent PROGRAM SUMMARY} \\
\begin{small}
\noindent
{\em Program Title:} PyStructureFactor \\
{\em CPC Library link to program files:} (to be added by Technical Editor) \\
{\em Developer's repository link:} \url{https://github.com/TheStarAlight/PyStructureFactor} \\
{\em Code Ocean capsule:} (to be added by Technical Editor) \\
{\em Licensing provisions:} Apache-2.0 \\
{\em Programming language:} Python 3                                  \\
{\em Supplementary material:}                                 \\
{\em Nature of problem:} The structure factor of a molecule in intense laser fields determines its orientation-dependent tunneling ionization rate, which is crucial for the studies of ultrafast molecular dynamics and its control. However, the strong-field community lacks an open-source code to calculate the molecular structure factor, and can only resort to known results of a few molecules. \\
{\em Solution method:} We developed the \texttt{PyStructureFactor} program with the structure factor of arbitrary molecules calculated using the weak-field asymptotic theory in the integral representation. The underlying electronic structure necessary for the calculation is obtained using the \texttt{PySCF} quantum chemistry package. \\
{\em Restrictions on the Accuracy:} The accuracy of the molecular structure factor calculated by \texttt{PyStructureFactor} is restricted by the level of precision of the electronic structure information extracted from the \texttt{PySCF} package. \\
\cyan{{\em Running time:} The running time depends on the size of the molecule, the basis set of the calculation, the level of precision of the electronic structure calculations, and other parameters passed to the program. The example in Fig.~\ref{fig:code_example} took 1.2 seconds to finish on an AMD Ryzen 9 7950X CPU on the WSL Ubuntu 22.04 LTS.}
\end{small}

@@

Advances in the laser technology have made intense laser fields widely accessible in typical table-top lab settings. When matter interacts with intense laser fields, a plethora of novel strong-field phenomena emerge that are of fundamental interest, such as high-order harmonic generation (HHG) \cite{krause_high_1992,corkum_plasma_1993,popmintchev_the_2010}, above-threshold ionization \cite{agostini_free_1979,becker_above_2002,milosevic_above_2006}, nonsequential double ionization \cite{walker_precision_1994}, and laser-induced electron diffraction \cite{blaga_imaging_2012,wolter_ultrafast_2016}. HHG is the key to \cyan{produce extreme ultraviolet light pulses}, which has facilitated the improvement of time resolution of ultraprecise measurements down to the attosecond level \cite{krausz_attosecond_2009,dahlstroem_introduction_2012,pazourek_attosecond_2015,kheifets_wigner_2023}. Common to many of these strong-field phenomena is their first step, known as tunneling ionization \cite{keldysh_ionization_1965,chin_from_2004}, which is thus a cornerstone of strong-field and attosecond science. It has attracted widespread interest not only in physics but also in chemistry and related fields \cyan{\cite{smirnova_attosecond_2009,sukiasyan_exchange_2010,xie_attosecond-recollision-controlled_2012,spanner_strong-field_2012,doblhoff-dier_classical_2013,doblhoff-dier_theoretical_2016,patchkovskii_full-dimensional_2017}}. Given the recent development in optochemistry \cite{li_light_2022,ma_transient_2021}, a complementary emerging field to photochemistry, the importance of the tunneling ionization of molecules has been elevated to new heights.

In contrast to photoionization where the \cyan{photonic feature of light plays a dominant role}, tunneling ionization is better represented by the highly nonlinear ionization induced by intense optical fields in the time domain $F(t)$, with a tunneling ionization rate \cite{ammosov_tunnel_1986,delone_tunneling_1998}
\begin{equation}
    W[F(t)] \sim \ee^{-2\kappa^3/3F(t)}
    \label{eq:FieldFactor}
\end{equation}
depending exponentially on the field strength $F(t)$, where $\kappa=\sqrt{-2E_0}$ with $E_0$ \cyan{the orbital energy of the target}. For typical intense infrared laser fields, the photon energy is substantially lower than the ionization potential of the bound electron. Moreover, the timescale of the field oscillation is much longer than the intrinsic attosecond timescale of the bound electron. Therefore, the strong-field ionization process can be approximated by tunneling in a static electric field represented by the instantaneous laser electric field $F(t)$ \cite{tolstikhin_adiabatic_2010, tolstikhin_adiabatic_2012}.

@@

The molecular structure factor can be obtained either by the molecular Ammosov--Delone--Krainov theory (MO-ADK) \cite{tong_theory_2002, pavicic_direct_2007, holmegaard_photoelectron_2010, johansen_alignment-dependent_2016, zhao_accurate_2017} or via the weak-field asymptotic theory (WFAT) \cite{tolstikhin_theory_2011, madsen_application_2012, madsen_structure_2013, trinh_first-order_2013, madsen_application_2014, tolstikhin_weak-field_2014, tolstikhina_application_2014, saito_structure_2015, trinh_weak-field_2015, dnestryan_integral-equation_2016, svensmark_theory_2016, trinh_first-order_2016, madsen_structure_2017, dnestryan_structure_2018, samygin_weak-field_2018, matsui_weak-field_2021}. WFAT is applicable in laser fields below the critical field strength that leads to over-barrier ionization. Based on the parabolic adiabatic expansion approach \cite{batishchev_atomic_2010}, WFAT generalizes isotropic atomic potentials to arbitrary molecular potentials and naturally accounts for the effect of molecular permanent dipole moment (if any), \cyan{and has been adopted as the theoretical method in some experimental research on molecular strong-field ionization \cite{ohmura_molecular_2014, kraus_observation_2015, walt_role_2015, endo_imaging_2016, fujise_helicity-dependent_2022}}. In this work, we employ WFAT towards a numerical implementation of the computation of the molecular structure factor. Under its framework, the obtained tunneling ionization rate, in its leading order, factorizes into the field factor [Eq.~(\ref{eq:FieldFactor})] and the absolute square of the structure factor, the latter of which is an inherent property of the ionizing molecular orbital independent of the external field \cite{trinh_first-order_2013}.

In an essence of the perturbation theory, the structure factor of WFAT is \cout{taken as the asymptotic expansion for the field strength} \cyan{found through the behavior of the wave function in the asymptotic region} in early stages, known as the tail representation \cite{madsen_structure_2013, trinh_weak-field_2015, madsen_application_2014}. Hence, its reliable evaluation requires an accurate description of the asymptotic tail of the ionizing orbital. To extract the orbital wave function of polyatomic molecules by the Hartree-Fock (HF) method or density functional theory (DFT) with quantum chemistry software packages \cite{madsen_application_2014, saito_structure_2015}, one normally resorts to the basis-based approaches. However, compared to the correct exponential behavior obtained by the grid-based approach, the standard Gaussian-type basis functions facilitating multicenter integral calculations decay too rapidly in the asymptotic region and thus could result in erratic oscillations. To reproduce the correct local property of the asymptotic tail of the orbital, large basis sets with quadruple or pentuple-zeta quality and variationally optimized exponents are required \cite{trinh_first-order_2013}. This approach significantly increases the computational cost and become infeasible in many cases. The efficiency of such numerical computation needs to be improved urgently.

@@

In this work, an open-source program implementing WFAT in the integral representation is developed that can easily and quickly calculate the molecular structure factor of common molecules. In our implementation, named the \texttt{PyStructureFactor}, the ionizing molecular orbital is expanded as a linear combination of GTOs at the HF \cyan{or complete-active-space self-consistent field (CASSCF)} level of theory calculated by the \texttt{PySCF} package \cite{sun_pyscf_2018, sun_recent_2020}. The integrals are evaluated numerically using a three-dimensional grid-based method, which exhibits much better numerical stability and is readily applicable to large molecules, compared to the standard analytical integration procedure. Our present implementation of \texttt{PyStructureFactor} features \cyan{flexible control of accuracy and} efficient computation towards the calculation of the molecular structure factor.

@@

Tunneling ionization of molecules can be usually modeled within the Born-Oppenheimer and the single-active-electron approximation. Under such a framework, the wave function of the ionizing orbital $\psi_0$ is the solution to an effective one-electron Schr\"{o}dinger equation
\begin{equation}
    \left[-\frac12\nabla^2 + \cyan{V}(\bn{r})\right] \psi_0(\bn{r}) = E_0 \psi_0(\bn{r}),
\end{equation}
where the HF potential $\cyan{V}(\bn{r})$ describes the interaction between the ionizing electron and the parent ion.

\begin{figure}[tb]
    \centering
    \includegraphics[width=0.5\columnwidth]{Euler_Angle_20230706.pdf}
    \caption{\cyan{
        Illustration of the rotation from the LF to the MF using Euler angles of $z$-$y'$-$z''$ convention. The rotation consists of three steps: (1) The coordinate system $(x,y,z)$ rotates about the $z$ axis by angle $\alpha$, then (2) rotates about the new $y$ axis by angle $\beta$, finally (3) rotates about the rotated $z$ axis by $\gamma$, and becomes $(X,Y,Z)$. The rotations mentioned above are all counter-clockwise.
        }}
    \label{fig:EulerAngle}
\end{figure}

Within WFAT, tunneling ionization is investigated in parabolic coordinates. We assume the electric field vector $\bn{F}=F\hat{\bn{z}}$, and define the orientation of the molecule using a set of Euler angles $(\alpha,\beta,\gamma)$ (within the $z$-$y'$-$z''$ convention \cite{madsen_structure_2017}, \cyan{which is shown schematically in Fig.~\ref{fig:EulerAngle}}) representing a rotation $\hat{\bn{R}}$ from the laboratory frame (LF) to the molecular frame (MF). The ionization rate does not depend on $\alpha$ and so we set $\alpha=0$ throughout. The total ionization rate is split into different parabolic channels:
\begin{equation}
    \Gamma(\beta, \gamma)=\sum_\nu \Gamma_\nu(\beta, \gamma),
\end{equation}
where $\Gamma_\nu(\beta, \gamma)$ are partial rates of parabolic quantum number indices
\begin{equation}
    \nu=(n_\xi,m), \quad n_\xi=0,1,2,\cdots, \quad m=0,\pm 1,\pm 2,\cdots
\end{equation}
In the leading-order approximation, the asymptotic expansion of the partial rates for $F\rightarrow 0$ gives
\begin{equation}
    \Gamma_{\nu}(\beta, \gamma) = \abs{G_\nu(\beta, \gamma)}^2 W_{\nu}(F),
\end{equation}
which can be separated into two factors: The structural part $\abs{G_\nu(\beta, \gamma)}^2$ and the field part $W_{\nu}(F)$.
The field factor, $W_{\nu}(F)$, has the form \cite{ammosov_tunnel_1986,delone_tunneling_1998}
\begin{equation}
    W_{\nu}(F) = \frac{\kappa}{2} \left(\frac{4\kappa^2}{F}\right)^{2Z/\kappa-2n_\xi-\abs{m}-1} \ee^{-{2\kappa^3}/{3F}}.
\end{equation}
The structure factor, namely $G_\nu(\beta, \gamma)$, in the integral representation \cyan{of WFAT}, is given as an integral:
\begin{equation}
    G_\nu (\beta,\gamma) = \ee^{-\kappa\mu_z} \int \Omega_\nu^* \left(\hat{\bn{R}}^{-1} \bn{r}\right) \cyan{V}_{\rm{c}}(\bn{r}) \psi_0(\bn{r}) \dd \bn{r},
    \label{eq:StructFactorExp}
\end{equation}
where
\begin{equation}
    \bi{\mu} = \int \psi_0^*(\bn{r}) \bn{r} \psi_0(\bn{r}) \dd \bn{r}
\end{equation}
denotes the orbital dipole moment in the LF with $\mu_z$ its component along the field direction;
$\Omega_\nu$ is a reference function which can be expanded into spherical harmonics:
\begin{equation}
    \Omega_\nu(\bn{r}) = \sum_{l=|m|}^{\infty} \Omega^\nu_{lm}(\bn{r}),
    \label{eq:OmegaExpansion1}
\end{equation}
where
\begin{equation}
    \Omega^\nu_{lm}(\bn{r}) = \sum_{l=|m|}^{\infty} R_l^\nu(r) Y_{lm}(\theta, \phi);
    \label{eq:OmegaExpansion2}
\end{equation}
$\cyan{V}_{\rm{c}}(\bn{r}) = \cyan{V}(\bn{r}) + Z/r$ is the core potential without the Coulomb tail (i.e., $\cyan{V}_{\rm{c}} \rightarrow 0$ as $r\rightarrow \infty$) with $Z$ the asymptotic charge of the parent molecular ion.
The radial part $R_l^\nu$ of the reference function $\Omega_{lm}^\nu$ in Eq.~(\ref{eq:OmegaExpansion2}) reads
\begin{equation}
    R_l^\nu(r)=\omega_l^\nu \ (\kappa r)^l \ \ee^{-\kappa r} \ \rm{M}(l+1-Z/\kappa, 2l+2, 2 \kappa r),
\end{equation}
where $\rm{M}(a, b, x)$ is the confluent hyper-geometric function \cite{olver_nist_2010} and $\omega_l^\nu$ is the normalization coefficient:
\begin{align}
    \omega_l^\nu = & \      (-1)^{l+(\abs{m}-m)/2+1}\ 2^{l+3/2}\ \kappa^{Z/\kappa-(\abs{m}+1)/2-n_\xi} \nonumber \\
                   & \times \sqrt{(2l+1)(l+m)!(l-m)!(\abs{m}+n_\xi)!n_\xi!}\ \frac{l!}{(2l+1)!} \nonumber \\
                   & \times \!\!\!\!\!\! \sum_{k=0}^{\min{(n_\xi,l-\abs{m})}} \!\!\!\!\!\!\!\!\!\! \frac{\Gamma(l+1-Z/\kappa+n_\xi-k)}{k!(l-k)!(\abs{m}+k)!(l-\abs{m}-k)!(n_\xi-k)!}.
\end{align}

Under the framework of the HF method, the HF potential $\cyan{V}(\bn{r})$ consists of three parts, namely the nuclear Coulomb potential ($V_{\rm{nuc}}$), the direct ($V_{\rm{d}}$) and exchange ($V_{\rm{ex}}$) parts of inter-electron interactions:
\begin{equation}
    \cyan{V}(\bn{r}) = V_{\rm{nuc}}(\bn{r}) + V_{\rm{d}}(\bn{r}) + \cyan{V}_{\rm{ex}}(\bn{r}),
\end{equation}
where
\begin{align}
    V_{\rm{nuc}}(\bn{r}) &= -\sum_{A=1}^{N_{\rm{atm}}} \frac{Z_A}{\left|\bn{r}-\bn{R}_A\right|}, \label{eq:Vnuc} \\
    V_{\rm{d}}(\bn{r}) &= \sum_{i=1}^N \int \frac{\psi_i^*(\bn{r}') \psi_i(\bn{r}')}{\abs{\bn{r}-\bn{r}'}} \dd \bn{r}', \label{eq:Vd} \\
    \cyan{V}_{\rm{ex}}(\bn{r}) \psi_0(\bn{r}) &= -\sum_{i=1}^N \psi_i(\bn{r}) \int \frac{\psi_i^*(\bn{r}') \psi_0(\bn{r}')}{\abs{\bn{r}-\bn{r}'}} \braket{\sigma_i | \sigma_0} \dd \bn{r}', \label{eq:VexPsi0}
\end{align}
where $N$ is the number of electrons, $N_{\rm{atm}}$ is the number of nuclei, $\psi_i(\bn{r})$ and $\sigma_i$ denote the molecular orbital and the spin state of the electron of index $i$ ($\braket{\sigma_i | \sigma_j}=1$ if electrons $i$ and $j$ have the same spin, and $\braket{\sigma_i | \sigma_j}=0$ otherwise), while $Z_A$ and $\bn{R}_A$ are the nuclear charge and position of atom of index $A$.

@@

For the practical numerical implementation of the computation of the molecular structure factor within the framework of WFAT, we first refer to the evaluation of the integral in Eq.~(\ref{eq:StructFactorExp}). Directly evaluating the integrals for each molecular orientation $(\beta,\gamma)$ is inefficient. Nevertheless, we may utilize the spherical harmonic expansion of the reference function $\Omega_\nu$, and express the rotated spherical harmonic functions as a linear combination of spherical harmonic functions using the Wigner-$d$ matrix. Substituting Eqs.~(\ref{eq:OmegaExpansion1}) and (\ref{eq:OmegaExpansion2}) into Eq.~(\ref{eq:StructFactorExp}) yields \cite{madsen_structure_2017}
\begin{equation}
    G_\nu(\beta,\gamma) = \ee^{-\kappa \mu_z} \sum_{l=|m|}^{\infty} \sum_{m'=-l}^l I_{l m'}^\nu  d_{m m'}^l(\beta) e^{-\ii m' \gamma},
    \label{eq:StructFactorExpInt}
\end{equation}
where $d_{m m'}^l(\beta)$ is the Wigner-$d$ matrix, and the integral $I_{l m'}^\nu$ is expressed as
\begin{equation}
    I_{l m'}^\nu = \int \Omega_{l m'}^{\nu *}(\bn{r}) \cyan{V}_{\rm{c}}(\bn{r}) \psi_0(\bn{r}) \dd \bn{r}.
    \label{eq:IntegralExpr}
\end{equation}
In this way the computational overhead is significantly reduced --- with the values of the integrals $I_{l m'}^\nu$, we may obtain the structure factor of arbitrary molecular orientation $G_\nu(\beta,\gamma)$ with little computational cost.

>>>>

\begin{abstract}
Tunneling ionization is at the core of strong-field and attosecond science. In this paper, we present \texttt{PyStructureFactor} --- a general \texttt{Python} code towards the calculation of the structure factor in the tunneling ionization rate of common molecules under intense laser fields. The numerical implementation is based on the well-developed weak-field asymptotic theory in the integral representation. The information of the electronic structure of the molecules is obtained via the \texttt{PySCF} quantum chemistry package. \texttt{PyStructureFactor} is a general computational framework that can be utilized to compute the molecular structure factor of various types of molecules, including polar and nonpolar diatomic molecules, degenerate molecules, and open-shell molecules. Examples are given that are benchmarked against known results with good agreements. The present \texttt{PyStructureFactor} is implemented in an efficient manner and is easily applicable towards larger molecules.
\end{abstract}

@@

{\bf \noindent PROGRAM SUMMARY} \\
\begin{small}
\noindent
{\em Program Title:} PyStructureFactor \\
{\em CPC Library link to program files:} (to be added by Technical Editor) \\
{\em Developer's repository link:} \url{https://github.com/TheStarAlight/PyStructureFactor} \\
{\em Code Ocean capsule:} (to be added by Technical Editor) \\
{\em Licensing provisions:} Apache-2.0 \\
{\em Programming language:} Python 3                                  \\
{\em Supplementary material:}                                 \\
{\em Nature of problem:} The structure factor of a molecule in intense laser fields determines its orientation-dependent tunneling ionization rate, which is crucial for the studies of ultrafast molecular dynamics and its control. However, the strong-field community lacks an open-source code to calculate the molecular structure factor, and can only resort to known results of a few molecules. \\
{\em Solution method:} We developed the \texttt{PyStructureFactor} program with the structure factor of arbitrary molecules calculated using the weak-field asymptotic theory in the integral representation. The underlying electronic structure necessary for the calculation is obtained using the \texttt{PySCF} quantum chemistry package. \\
{\em Restrictions on the Accuracy:} The accuracy of the molecular structure factor calculated by \texttt{PyStructureFactor} is restricted by the level of precision of the electronic structure information extracted from the \texttt{PySCF} package. \\
{\em Running time:} The running time depends on the size of the molecule, the basis set of the calculation, the level of precision of the electronic structure calculations, and other parameters passed to the program. The example in Fig.~\ref{fig:code_example} took 1.2 seconds to finish on an AMD Ryzen 9 7950X CPU on the WSL Ubuntu 22.04 LTS.
\end{small}

@@

Advances in the laser technology have made intense laser fields widely accessible in typical table-top lab settings. When matter interacts with intense laser fields, a plethora of novel strong-field phenomena emerge that are of fundamental interest, such as high-order harmonic generation (HHG) \cite{krause_high_1992,corkum_plasma_1993,popmintchev_the_2010}, above-threshold ionization \cite{agostini_free_1979,becker_above_2002,milosevic_above_2006}, nonsequential double ionization \cite{walker_precision_1994}, and laser-induced electron diffraction \cite{blaga_imaging_2012,wolter_ultrafast_2016}. HHG is the key to produce extreme ultraviolet light pulses, which has facilitated the improvement of time resolution of ultraprecise measurements down to the attosecond level \cite{krausz_attosecond_2009,dahlstroem_introduction_2012,pazourek_attosecond_2015,kheifets_wigner_2023}. Common to many of these strong-field phenomena is their first step, known as tunneling ionization \cite{keldysh_ionization_1965,chin_from_2004}, which is thus a cornerstone of strong-field and attosecond science. It has attracted widespread interest not only in physics but also in chemistry and related fields \cite{smirnova_attosecond_2009,sukiasyan_exchange_2010,xie_attosecond-recollision-controlled_2012,spanner_strong-field_2012,doblhoff-dier_classical_2013,doblhoff-dier_theoretical_2016,patchkovskii_full-dimensional_2017}. Given the recent development in optochemistry \cite{li_light_2022,ma_transient_2021}, a complementary emerging field to photochemistry, the importance of the tunneling ionization of molecules has been elevated to new heights.

In contrast to photoionization where the photonic feature of light plays a dominant role, tunneling ionization is better represented by the highly nonlinear ionization induced by intense optical fields in the time domain $F(t)$, with a tunneling ionization rate \cite{ammosov_tunnel_1986,delone_tunneling_1998}
\begin{equation}
    W[F(t)] \sim \ee^{-2\kappa^3/3F(t)}
    \label{eq:FieldFactor}
\end{equation}
depending exponentially on the field strength $F(t)$, where $\kappa=\sqrt{-2E_0}$ with $E_0$ the orbital energy of the target. For typical intense infrared laser fields, the photon energy is substantially lower than the ionization potential of the bound electron. Moreover, the timescale of the field oscillation is much longer than the intrinsic attosecond timescale of the bound electron. Therefore, the strong-field ionization process can be approximated by tunneling in a static electric field represented by the instantaneous laser electric field $F(t)$ \cite{tolstikhin_adiabatic_2010, tolstikhin_adiabatic_2012}.

@@

The molecular structure factor can be obtained either by the molecular Ammosov--Delone--Krainov theory (MO-ADK) \cite{tong_theory_2002, pavicic_direct_2007, holmegaard_photoelectron_2010, johansen_alignment-dependent_2016, zhao_accurate_2017} or via the weak-field asymptotic theory (WFAT) \cite{tolstikhin_theory_2011, madsen_application_2012, madsen_structure_2013, trinh_first-order_2013, madsen_application_2014, tolstikhin_weak-field_2014, tolstikhina_application_2014, saito_structure_2015, trinh_weak-field_2015, dnestryan_integral-equation_2016, svensmark_theory_2016, trinh_first-order_2016, madsen_structure_2017, dnestryan_structure_2018, samygin_weak-field_2018, matsui_weak-field_2021}. WFAT is applicable in laser fields below the critical field strength that leads to over-barrier ionization. Based on the parabolic adiabatic expansion approach \cite{batishchev_atomic_2010}, WFAT generalizes isotropic atomic potentials to arbitrary molecular potentials and naturally accounts for the effect of molecular permanent dipole moment (if any), and has been adopted as the theoretical method in some experimental research on molecular strong-field ionization \cite{ohmura_molecular_2014, kraus_observation_2015, walt_role_2015, endo_imaging_2016, fujise_helicity-dependent_2022}. In this work, we employ WFAT towards a numerical implementation of the computation of the molecular structure factor. Under its framework, the obtained tunneling ionization rate, in its leading order, factorizes into the field factor [Eq.~(\ref{eq:FieldFactor})] and the absolute square of the structure factor, the latter of which is an inherent property of the ionizing molecular orbital independent of the external field \cite{trinh_first-order_2013}.

In an essence of the perturbation theory, the structure factor of WFAT is found through the behavior of the wave function in the asymptotic region in early stages, known as the tail representation \cite{madsen_structure_2013, trinh_weak-field_2015, madsen_application_2014}. Hence, its reliable evaluation requires an accurate description of the asymptotic tail of the ionizing orbital. To extract the orbital wave function of polyatomic molecules by the Hartree-Fock (HF) method or density functional theory (DFT) with quantum chemistry software packages \cite{madsen_application_2014, saito_structure_2015}, one normally resorts to the basis-based approaches. However, compared to the correct exponential behavior obtained by the grid-based approach, the standard Gaussian-type basis functions facilitating multicenter integral calculations decay too rapidly in the asymptotic region and thus could result in erratic oscillations. To reproduce the correct local property of the asymptotic tail of the orbital, large basis sets with quadruple or pentuple-zeta quality and variationally optimized exponents are required \cite{trinh_first-order_2013}. This approach significantly increases the computational cost and become infeasible in many cases. The efficiency of such numerical computation needs to be improved urgently.

@@

In this work, an open-source program implementing WFAT in the integral representation is developed that can easily and quickly calculate the molecular structure factor of common molecules. In our implementation, named the \texttt{PyStructureFactor}, the ionizing molecular orbital is expanded as a linear combination of GTOs at the HF or complete-active-space self-consistent field (CASSCF) level of theory calculated by the \texttt{PySCF} package \cite{sun_pyscf_2018, sun_recent_2020}. The integrals are evaluated numerically using a three-dimensional grid-based method, which exhibits much better numerical stability and is readily applicable to large molecules, compared to the standard analytical integration procedure. Our present implementation of \texttt{PyStructureFactor} features flexible control of accuracy and efficient computation towards the calculation of the molecular structure factor.

@@

Tunneling ionization of molecules can be usually modeled within the Born-Oppenheimer and the single-active-electron approximation. Under such a framework, the wave function of the ionizing orbital $\psi_0$ is the solution to an effective one-electron Schr\"{o}dinger equation
\begin{equation}
    \left[-\frac12\nabla^2 + V(\bn{r})\right] \psi_0(\bn{r}) = E_0 \psi_0(\bn{r}),
\end{equation}
where the HF potential $V(\bn{r})$ describes the interaction between the ionizing electron and the parent ion.

\begin{figure}[tb]
    \centering
    \includegraphics[width=0.5\columnwidth]{Euler_Angle_20230706.pdf}
    \caption{
        Illustration of the rotation from the LF to the MF using Euler angles of $z$-$y'$-$z''$ convention. The rotation consists of three steps: (1) The coordinate system $(x,y,z)$ rotates about the $z$ axis by angle $\alpha$, then (2) rotates about the new $y$ axis by angle $\beta$, finally (3) rotates about the rotated $z$ axis by $\gamma$, and becomes $(X,Y,Z)$. The rotations mentioned above are all counter-clockwise.
        }
    \label{fig:EulerAngle}
\end{figure}

Within WFAT, tunneling ionization is investigated in parabolic coordinates. We assume the electric field vector $\bn{F}=F\hat{\bn{z}}$, and define the orientation of the molecule using a set of Euler angles $(\alpha,\beta,\gamma)$ (within the $z$-$y'$-$z''$ convention \cite{madsen_structure_2017}, which is shown schematically in Fig.~\ref{fig:EulerAngle}) representing a rotation $\hat{\bn{R}}$ from the laboratory frame (LF) to the molecular frame (MF). The ionization rate does not depend on $\alpha$ and so we set $\alpha=0$ throughout. The total ionization rate is split into different parabolic channels:
\begin{equation}
    \Gamma(\beta, \gamma)=\sum_\nu \Gamma_\nu(\beta, \gamma),
\end{equation}
where $\Gamma_\nu(\beta, \gamma)$ are partial rates of parabolic quantum number indices
\begin{equation}
    \nu=(n_\xi,m), \quad n_\xi=0,1,2,\cdots, \quad m=0,\pm 1,\pm 2,\cdots
\end{equation}
In the leading-order approximation, the asymptotic expansion of the partial rates for $F\rightarrow 0$ gives
\begin{equation}
    \Gamma_{\nu}(\beta, \gamma) = \abs{G_\nu(\beta, \gamma)}^2 W_{\nu}(F),
\end{equation}
which can be separated into two factors: The structural part $\abs{G_\nu(\beta, \gamma)}^2$ and the field part $W_{\nu}(F)$.
The field factor, $W_{\nu}(F)$, has the form \cite{ammosov_tunnel_1986,delone_tunneling_1998}
\begin{equation}
    W_{\nu}(F) = \frac{\kappa}{2} \left(\frac{4\kappa^2}{F}\right)^{2Z/\kappa-2n_\xi-\abs{m}-1} \ee^{-{2\kappa^3}/{3F}}.
\end{equation}
The structure factor, namely $G_\nu(\beta, \gamma)$, in the integral representation of WFAT, is given as an integral:
\begin{equation}
    G_\nu (\beta,\gamma) = \ee^{-\kappa\mu_z} \int \Omega_\nu^* \left(\hat{\bn{R}}^{-1} \bn{r}\right) V_{\rm{c}}(\bn{r}) \psi_0(\bn{r}) \dd \bn{r},
    \label{eq:StructFactorExp}
\end{equation}
where
\begin{equation}
    \bi{\mu} = \int \psi_0^*(\bn{r}) \bn{r} \psi_0(\bn{r}) \dd \bn{r}
\end{equation}
denotes the orbital dipole moment in the LF with $\mu_z$ its component along the field direction;
$\Omega_\nu$ is a reference function which can be expanded into spherical harmonics:
\begin{equation}
    \Omega_\nu(\bn{r}) = \sum_{l=|m|}^{\infty} \Omega^\nu_{lm}(\bn{r}),
    \label{eq:OmegaExpansion1}
\end{equation}
where
\begin{equation}
    \Omega^\nu_{lm}(\bn{r}) = \sum_{l=|m|}^{\infty} R_l^\nu(r) Y_{lm}(\theta, \phi);
    \label{eq:OmegaExpansion2}
\end{equation}
$V_{\rm{c}}(\bn{r}) = V(\bn{r}) + Z/r$ is the core potential without the Coulomb tail (i.e., $V_{\rm{c}} \rightarrow 0$ as $r\rightarrow \infty$) with $Z$ the asymptotic charge of the parent molecular ion.
The radial part $R_l^\nu$ of the reference function $\Omega_{lm}^\nu$ in Eq.~(\ref{eq:OmegaExpansion2}) reads
\begin{equation}
    R_l^\nu(r)=\omega_l^\nu \ (\kappa r)^l \ \ee^{-\kappa r} \ \rm{M}(l+1-Z/\kappa, 2l+2, 2 \kappa r),
\end{equation}
where $\rm{M}(a, b, x)$ is the confluent hyper-geometric function \cite{olver_nist_2010} and $\omega_l^\nu$ is the normalization coefficient:
\begin{align}
    \omega_l^\nu = & \      (-1)^{l+(\abs{m}-m)/2+1}\ 2^{l+3/2}\ \kappa^{Z/\kappa-(\abs{m}+1)/2-n_\xi} \nonumber \\
                   & \times \sqrt{(2l+1)(l+m)!(l-m)!(\abs{m}+n_\xi)!n_\xi!}\ \frac{l!}{(2l+1)!} \nonumber \\
                   & \times \!\!\!\!\!\! \sum_{k=0}^{\min{(n_\xi,l-\abs{m})}} \!\!\!\!\!\!\!\!\!\! \frac{\Gamma(l+1-Z/\kappa+n_\xi-k)}{k!(l-k)!(\abs{m}+k)!(l-\abs{m}-k)!(n_\xi-k)!}.
\end{align}

Under the framework of the HF method, the HF potential $V(\bn{r})$ consists of three parts, namely the nuclear Coulomb potential ($V_{\rm{nuc}}$), the direct ($V_{\rm{d}}$) and exchange ($V_{\rm{ex}}$) parts of inter-electron interactions:
\begin{equation}
    V(\bn{r}) = V_{\rm{nuc}}(\bn{r}) + V_{\rm{d}}(\bn{r}) + V_{\rm{ex}}(\bn{r}),
\end{equation}
where
\begin{align}
    V_{\rm{nuc}}(\bn{r}) &= -\sum_{A=1}^{N_{\rm{atm}}} \frac{Z_A}{\left|\bn{r}-\bn{R}_A\right|}, \label{eq:Vnuc} \\
    V_{\rm{d}}(\bn{r}) &= \sum_{i=1}^N \int \frac{\psi_i^*(\bn{r}') \psi_i(\bn{r}')}{\abs{\bn{r}-\bn{r}'}} \dd \bn{r}', \label{eq:Vd} \\
    V_{\rm{ex}}(\bn{r}) \psi_0(\bn{r}) &= -\sum_{i=1}^N \psi_i(\bn{r}) \int \frac{\psi_i^*(\bn{r}') \psi_0(\bn{r}')}{\abs{\bn{r}-\bn{r}'}} \braket{\sigma_i | \sigma_0} \dd \bn{r}', \label{eq:VexPsi0}
\end{align}
where $N$ is the number of electrons, $N_{\rm{atm}}$ is the number of nuclei, $\psi_i(\bn{r})$ and $\sigma_i$ denote the molecular orbital and the spin state of the electron of index $i$ ($\braket{\sigma_i | \sigma_j}=1$ if electrons $i$ and $j$ have the same spin, and $\braket{\sigma_i | \sigma_j}=0$ otherwise), while $Z_A$ and $\bn{R}_A$ are the nuclear charge and position of atom of index $A$.

@@

For the practical numerical implementation of the computation of the molecular structure factor within the framework of WFAT, we first refer to the evaluation of the integral in Eq.~(\ref{eq:StructFactorExp}). Directly evaluating the integrals for each molecular orientation $(\beta,\gamma)$ is inefficient. Nevertheless, we may utilize the spherical harmonic expansion of the reference function $\Omega_\nu$, and express the rotated spherical harmonic functions as a linear combination of spherical harmonic functions using the Wigner-$d$ matrix. Substituting Eqs.~(\ref{eq:OmegaExpansion1}) and (\ref{eq:OmegaExpansion2}) into Eq.~(\ref{eq:StructFactorExp}) yields \cite{madsen_structure_2017}
\begin{equation}
    G_\nu(\beta,\gamma) = \ee^{-\kappa \mu_z} \sum_{l=|m|}^{\infty} \sum_{m'=-l}^l I_{l m'}^\nu  d_{m m'}^l(\beta) e^{-\ii m' \gamma},
    \label{eq:StructFactorExpInt}
\end{equation}
where $d_{m m'}^l(\beta)$ is the Wigner-$d$ matrix, and the integral $I_{l m'}^\nu$ is expressed as
\begin{equation}
    I_{l m'}^\nu = \int \Omega_{l m'}^{\nu *}(\bn{r}) V_{\rm{c}}(\bn{r}) \psi_0(\bn{r}) \dd \bn{r}.
    \label{eq:IntegralExpr}
\end{equation}
In this way the computational overhead is significantly reduced --- with the values of the integrals $I_{l m'}^\nu$, we may obtain the structure factor of arbitrary molecular orientation $G_\nu(\beta,\gamma)$ with little computational cost.
